Based on polling as of 27 October 2024, Harris expects to win on 49.2%, leads Trump by 2% in 2024 US election

### 2024-10-30 paper/paper.qmd
update paper results

library(tidyverse)
library(arrow)
```

@@

```{r}
#| include: false
#| warning: false
#| message: false
#### Calculate Weights ####
# Calculate weights for Harris data
harris_data <- harris_data %>%
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) %>%
  group_by(pollster) %>%
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) %>%
  ungroup() %>%
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

# Calculate weights for Trump data
trump_data <- trump_data %>%
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) %>%
  group_by(pollster) %>%
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) %>%
  ungroup() %>%
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)
```

@@

harris_data <- harris_data %>%
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) %>%
  group_by(pollster) %>%
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) %>%
  ungroup() %>%
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

# Calculate weights for Trump data
trump_data <- trump_data %>%
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) %>%
  group_by(pollster) %>%
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) %>%
  ungroup() %>%
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

@@

harris_filtered <- harris_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)|>mutate(candidate = "Harris")
trump_filtered <- trump_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)|>mutate(candidate = "Trump")

combined_data <- bind_rows(harris_filtered, trump_filtered)
# Plot with recent data, specific axis formatting, and smoother trend lines
ggplot(combined_data, aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  facet_wrap(~ state) +
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Selected States",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-31"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))
```
@fig-support-trends-by-state illustrates the recent support trends for Kamala Harris and Donald Trump across selected battleground states from August to November 2024. Each subplot represents a different state, with support percentages for each candidate displayed over time. Key features include Harris shown in blue and Trump in red, with smoothed trend lines to capture changes over time. The plot's y-axis spans from approximately 30% to 50% support, while the x-axis highlights monthly intervals from August to November. For example, in Arizona, Trump maintains a slight lead with support fluctuating around 50% in October, compared to Harris around 48%. In Georgia, Trump's support reaches just above 49% in late September, whereas Harris hovers closer to 47%. In Wisconsin, both candidates show tightly aligned trends near 48%, highlighting a close race. Each point's size reflects the poll weight—calculated using recency, pollster quality, and sample size—indicating the relative importance of each poll. This visualization emphasizes recent trends in candidate support across key battleground states.

@@

>>>>

library(tidyverse)
library(arrow)
library(dplyr)
```

@@

```{r}
#| include: false
#| warning: false
#| message: false
#### Calculate Weights ####
# Calculate weights for Harris data
harris_data <- harris_data |>
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) |>
  group_by(pollster) |>
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>
  ungroup() |>
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

# Calculate weights for Trump data
trump_data <- trump_data |>
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) |>
  group_by(pollster) |>
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>
  ungroup() |>
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)
```

@@

harris_data <- harris_data |>
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) |>
  group_by(pollster) |>
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>
  ungroup() |>
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

# Calculate weights for Trump data
trump_data <- trump_data |>
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) |>
  group_by(pollster) |>
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>
  ungroup() |>
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

@@

harris_filtered <- harris_data |>
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)|>mutate(candidate = "Harris")
trump_filtered <- trump_data |>
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)|>mutate(candidate = "Trump")

combined_data <- bind_rows(harris_filtered, trump_filtered)
# Plot with recent data, specific axis formatting, and smoother trend lines
ggplot(combined_data, aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  facet_wrap(~ state) +
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Selected States",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-31"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(theme(panel.spacing = unit(1, "lines"))
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))
```
@fig-support-trends-by-state illustrates the recent support trends for Kamala Harris and Donald Trump across selected battleground states from August to November 2024. Each subplot represents a different state, with support percentages for each candidate displayed over time. Key features include Harris shown in blue and Trump in red, with smoothed trend lines to capture changes over time. The plot's y-axis spans from approximately 30% to 50% support, while the x-axis highlights monthly intervals from August to November. For example, in Arizona, Trump maintains a slight lead with support fluctuating around 50% in October, compared to Harris around 48%. In Georgia, Trump's support reaches just above 49% in late September, whereas Harris hovers closer to 47%. In Wisconsin, both candidates show tightly aligned trends near 48%, highlighting a close race. Each point's size reflects the poll weight—calculated using recency, pollster quality, and sample size—indicating the relative importance of each poll. This visualization emphasizes recent trends in candidate support across key battleground states.

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 1000 
  ),
  pollster = factor("YouGov", levels = levels(harris_data$pollster)),  # Ensure it matches training levels
  national_poll = rep(c(0, 1), length(1000)),  # Create a sequence for national_poll from 0 to 1
  population = factor("lv", levels = levels(harris_data$population)),  # Use the mean population or any relevant value
  sample_size = seq(min(0),max(2300),length.out=1000)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = 2 / 4,  # Example; replace with actual numeric grade if available

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime(as.Date("2024-11-05"), as.Date("2024-08-01") + (1:1000), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed

    # Poll frequency weight (you can set a condition based on a hypothetical recent poll count)
    poll_frequency_weight = ifelse(national_poll > 0.5, 1, 0.5),  # Example logic for poll frequency

    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight
  )

new_data_trump <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 1000
  ),
  pollster = factor("YouGov", levels = levels(harris_data$pollster)),  # Ensure it matches training levels
  national_poll = rep(c(0, 1), length(1000)),  # Create a sequence for national_poll from 0 to 1
  population = factor("lv", levels = levels(harris_data$population)),  # Use the mean population or any relevant value
  sample_size = seq(min(0),max(2300),length.out=1000)
)|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = 2 / 4,  # Example; replace with actual numeric grade if available

    # Recency weight (this can be based on the date difference from a certain cutoff, e.g., "2024-11-05")
    recency = as.numeric(difftime(as.Date("2024-11-05"), as.Date("2024-08-01") + (1:1000), units = "days")),  # Adjust for the sequence
    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed

    # Poll frequency weight (you can set a condition based on a hypothetical recent poll count)
    poll_frequency_weight = ifelse(national_poll > 0.5, 1, 0.5),  # Example logic for poll frequency

    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight
  )
```

```{r}
# Predictions for Harris
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(model_harris_unweighted, newdata = new_data_harris),
    predicted_weighted = predict(model_harris_weighted, newdata = new_data_harris)
  )

# Predictions for Trump
new_data_trump <- new_data_trump |>
  mutate(
    predicted_unweighted = predict(model_trump_unweighted, newdata = new_data_trump),
    predicted_weighted = predict(model_trump_weighted, newdata = new_data_trump)
  )

# Plot the predictions
library(dplyr)
library(ggplot2)

# Define the cutoff date for recent data
recent_date_cutoff <- as.Date("2024-05-01")

# Filter combined_predictions to include only data from the recent date cutoff onward
combined_predictions_filtered <- combined_predictions %>%
  filter(end_date >= recent_date_cutoff)

# Determine the maximum end date for setting the x-axis limit
max_end_date <- max(combined_predictions_filtered$end_date)

# Plotting with the filtered data
ggplot(combined_predictions_filtered, aes(x = end_date, color = candidate)) +
  geom_smooth(aes(y = predicted_unweighted, linetype = "Unweighted"), 
              method = "loess", se = FALSE, size = 1) +  # Unweighted model with smoothing
  geom_smooth(aes(y = predicted_weighted, linetype = "Weighted"), 
              method = "loess", se = FALSE, size = 1) +  # Weighted model with smoothing
  labs(title = "Predicted Support for Kamala Harris and Donald Trump \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_linetype_manual(name = "Linear Model Type", values = c("Unweighted" = "dashed", "Weighted" = "solid")) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

```
The predicted support for Kamala Harris and Donald Trump, as derived from multiple linear regression models, is displayed in the graph above. The analysis spans from August 1, 2024, to October 27, 2024, and highlights both unweighted and weighted predictions for each candidate.

For Kamala Harris, the unweighted model predicts her support to remain relatively stable, maintaining an approximate support level of 49%. This trend is visually represented by the blue dashed line, which indicates minimal fluctuations over time. Conversely, the weighted model shows a more pronounced trend, demonstrating a gradual increase in predicted support, suggesting a potential shift in voter sentiment towards Harris as the election date approaches.

In contrast, Donald Trump's predicted support, represented by the red lines, exhibits a downward trend in both unweighted and weighted models. The unweighted model indicates that his support is expected to hover around 48%, while the weighted model predicts a slight decline to approximately 46%. This trend may reflect challenges in maintaining voter enthusiasm in the face of changing public sentiment.

At the end of the polling period, we predict that Kamala Harris leads Donald Trump by 2% in the unweighted model (48% vs. 46%) and by about 2% in the weighted model (49.1% vs. 47.3%).
